### 2016-11-25 workspace/DoubleBars/MScProjectDoubleBars.tex

\documentclass[a4paper,12pt]{article}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{float}
\usepackage{natbib}
\usepackage{listings}
\graphicspath{{./}}
\DeclareGraphicsExtensions{.eps}
\author{Howard Kinsman}
\title{MSc Project - Modelling a Double Barred Galaxy as a Double Binary}
\begin{document}
\maketitle
\section{Introduction}
The aim of this project was to create a simple model of a double barred galaxy using a double binary.

\section{Methods}
Fortran 90 was selected as the language of choice in order to avoid the fixed line limitations of Fortran 77.
The model was created using OdeInt from Numerical Recipes. I first converted OdeInt to Fortran 90 for this purpose.
\newline
The classical equations of motion for an n-body problem are:
\begin{equation}
m_i\mathbf{\ddot{r}}=\sum_{i\neq{j}}\frac{{m_i}{m_j}}{r^3_{ij}}\mathbf{r_{ij}}
\qquad
i=1,2,3...
\end{equation}
where $\mathbf{r_i}=\left(x_i,y_j\right)$ and $\mathbf{r_{ij}}=\mathbf{r_j}-\mathbf{r_i}$. So for a four body problem 
this equation results in the following:
\begin{equation}
\ddot{\mathbf{r_1}}=\frac{m_2\mathbf{r_{12}}}{r^3_{12}}+\frac{m_3\mathbf{r_{13}}}{r^3_{13}}+\frac{m_4\mathbf{r_{14}}}{r^3_{14}}
\end{equation}
\begin{equation}
\ddot{\mathbf{r_2}}=\frac{m_1\mathbf{r_{21}}}{r^3_{21}}+\frac{m_3\mathbf{r_{23}}}{r^3_{23}}+\frac{m_4\mathbf{r_{24}}}{r^3_{24}}
\end{equation}
\begin{equation}
\ddot{\mathbf{r_3}}=\frac{m_1\mathbf{r_{31}}}{r^3_{31}}+\frac{m_2\mathbf{r_{32}}}{r^3_{32}}+\frac{m_4\mathbf{r_{34}}}{r^3_{34}}
\end{equation}
\begin{equation}
\ddot{\mathbf{r_4}}=\frac{m_1\mathbf{r_{41}}}{r^3_{41}}+\frac{m_2\mathbf{r_{42}}}{r^3_{42}}+\frac{m_3\mathbf{r_{43}}}{r^3_{43}}
\end{equation}
The above equations were encoded into Fortran. A full code listing is supplied in a separate file.
\subsection{Initial Conditions}
As a starting point both the outer binary and inner binary were placed in circular orbits and tested separately and then combined to ensure a stable
starting point. The inner binary had initially a neglible mass.
The outer binary initial conditions were:
\begin{lstlisting}
m1=.5, m2=.5, x1=-.5, x2=.5, y1=0, y2=0,
vx1=0, vx2=0, vy1=-.5, vy2=.5
\end{lstlisting}
and for the inner binary:
\begin{lstlisting}
m3=.001, m3=.001, x3=-.001, x4=.001, y3=0, y4=0, 
vx3=0, vx4=0, vy3=-.5, vy4=.5
\end{lstlisting}
A plot of this initial configuration is shown in Fig. \ref{fig:stablebase}.
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/stablebase/Orbit.eps}
\caption{Initial stable configuration}
\label{fig:stablebase}
\end{figure}
The inner binary is only visible as a point at the centre of the outer binary.
I then proceeded to perturb the system by gradually increasing the mass of the inner binary. As the system became unstable I componensated for
the increasing mass of the inner binary by increasing the velocity of the outer binary and also decreasing the velocity and increasing the binary 
separation of the inner binary. In many of the configurations the systems were inherently unstable and produced collisions with some 
or all of the bodies being ejected from the system. These were rejected and only 'stable' configurations were considered.
Table \ref{tab:variables} shows the perturbations made to the system - only the variables in this table changed and all the other initial
conditions remained the same.
\begin{table}[ht!]
  \centering
  \caption{Configurations}
  \label{tab:variables}
  \begin{tabular}{ccccccccc}
   Config. & m3 & m4 & vy1 & vy2 & x3 & x4 & vy3 & vy4\\
    \hline
   1 & .002 & .002 & -.5 & .5 & -.001 & .001 & -.5 & .5\\
   2 & .003 & .003 & -.5 & .5 & -.001 & .001 & -.5 & .5\\
   3 & .004 & .004 & -.5 & .5 & -.004 & .004 & -.5 & .5\\
   4 & .004 & .004 & -.55 & .55 & -.004 & .004 & -.5 & .5\\
   5 & .004 & .004 & -.57 & .57 & -.004 & .004 & -.5 & .5\\
   6 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.3 & .3\\
   7 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.4 & .4\\
   8 & .005 & .005 & -.58 & .58 & -.005 & .005 & -.35 & .35\\
   9 & .006 & .006 & -.6 & .6 & -.006 & .006 & -.3 & .3\\
   10 & .01 & .01 & -.7 & .7 & -.01 & .01 & -.2 & .2\\
   11 & .02 & .02 & -.75 & .75 & -.02 & .02 & -.1 & .1\\
   12 & .025 & .025 & -.65 & .65 & -.02 & .02 & -.5 & .5\\
   13 & .05 & .05 & -.9 & .9 & -.05 & .05 & -.1 & .1\\
   14 & .05 & .05 & -.9 & .9 & -.05 & .05 & -.15 & .15\\
   15 & .06 & .06 & -.95 & .95 & -.06 & .06 & -.15 & .15\\
   16 & .07 & .07 & -.1 & .1 & -.07 & .07 & -.15 & .15\\
   17 & .08 & .08 & -.105 & .105 & -.08 & .08 & -.12 & .12\\
  \end{tabular}
\end{table}

\section{Results}
As can be seen from Fig. \ref{fig:stablebase} even a neglible mass inner binary causes the outer binary to deviate from a circular orbit.
Figures Fig. \ref{fig:config1} to \ref{fig:config17} show plots of the double binary using the parameters given in Table \ref{tab:variables}.
As the mass of the inner binary increased the system became increasingly unstable and the effect on the outer binary became more pronounced.
Even the 'stable' configurations listed in Table \ref{tab:variables} hold for only a few orbits.
I was unable to recreate a stable system for a mass $>.08$ for the inner bar.
The plots below reveal that increasing the mass of the inner binary causes the outer binary to separate and
form wider orbits.

\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/002-5-001/Orbit.eps}
\caption{Configuration 1}
\label{fig:config1}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/003-5-001/Orbit.eps}
\caption{Configuration 2}
\label{fig:config2}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-5-004/Orbit.eps}
\caption{Configuration 3}
\label{fig:config3}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-55-004/Orbit.eps}
\caption{Configuration 4}
\label{fig:config4}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/004-57-004/Orbit.eps}
\caption{Configuration 5}
\label{fig:config5}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-3/Orbit.eps}
\caption{Configuration 6}
\label{fig:config6}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-4/Orbit.eps}
\caption{Configuration 7}
\label{fig:config7}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/005-58-005-35/Orbit.eps}
\caption{Configuration 8}
\label{fig:config8}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/006-6-006-3/Orbit.eps}
\caption{Configuration 9}
\label{fig:config9}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/01-7-01-2/Orbit.eps}
\caption{Configuration 10}
\label{fig:config10}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/02-75-02-1/Orbit.eps}
\caption{Configuration 11}
\label{fig:config11}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/025-65-02/Orbit.eps}
\caption{Configuration 12}
\label{fig:config12}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/05-9-05-1/Orbit.eps}
\caption{Configuration 13}
\label{fig:config13}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/05-9-05-15/Orbit.eps}
\caption{Configuration 14}
\label{fig:config14}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/06-95-06-15/Orbit.eps}
\caption{Configuration 15}
\label{fig:config15}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/07-1-07-15/Orbit.eps}
\caption{Configuration 16}
\label{fig:config16}
\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.75\textwidth]{./results/08-105-08-12/Orbit.eps}
\caption{Configuration 17}
\label{fig:config17}
\end{figure}
\end{document}